Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-14 ICML-submission/from-linear-separability-to-strong-linear-separability.tex
Second pass over section 5

Using the lemma and the polynomial approximation theorems, we can prove
the mappings $\phi_n, \psi_n, \rho$ map any a weakly separable data
to a strongly separable data set.

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T) \in \R^d \times \{1,2,\dots,K\}$
labeled examples that are weakly linearly separable with margin $\gamma > 0$
such that $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le 1$. Then, the labeled
examples $(\rho(x_1), y_1), (\rho(x_2), y_2), \dots, (\rho(x_T), y_T)$ are
strongly linearly separable with margin
$$
\gamma' = \frac{1}{2\sqrt{K} \left(376 \lceil \log_2(2m) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2m) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}}
$$
and $\norm{\rho(x_1)}_{\ell_2}, \norm{\rho(x_2)}_{\ell_2}, \dots, \norm{\rho(x_T)}_{\ell_2} \le 2$.
TODO
\end{theorem}

\begin{proof}
First osberve that for any $t=1,2,\dots,T$,
$$
\norm{\rho(x_t)}_{\ell_2} = K_{\rho}(x_t,x_t) = \frac{1}{1 - \frac{1}{2} \norm{x_t}^2} \le 2 \; .
$$

@@

Fix any $i \in \{1,2,\dots,K\}$. Consider the $K-1$ vectors
$(w_i - w_j)/2$ for $j \in \{1,2,\dots,K\} \setminus \{i\}$. Note that the vectors
have norm at most $1$.
\autoref{theorem:polynomial-approximation-1} implies that
there exists a polynomial $p_i:\R^d \to \R$ such
that
$$
\deg(p) = \lceil \log_2(2m) \rceil \cdot \lceil \sqrt{\frac{2}{\gamma}} \rceil \; ,
$$
for all $t=1,2,\dots,T$, if $y_t = i$ then $p_i(x_t) \ge 1/2$ and if $y_t \neq i$ then $p_i(x_t) \le -1/2$,
and
$$
\norm{p} \le \left(188 \lceil \log_2(2m) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2m) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \; .
$$
By \autoref{lemma:norm-bound} there exists $c_i \in \ell_2$ such that
$\ip{c_i}{\rho(x)} = p(x)$ and
$$
\norm{c_i}_{\ell_2}
\le \left(376 \lceil \log_2(2m) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2m) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \; .
$$
Define vector $u_i \in \ell_2$ as
$$
u_i = \frac{c_i}{\sqrt{K} \left(376 \lceil \log_2(2m) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2m) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}} \; .
$$

Then, $\norm{u_1}^2 + \norm{u_2}^2 + \dots + \norm{u_K}^2 \le 1$ and
for all $t=1,2,\dots,T$, $\ip{u_{y_t}}{x_t} \ge \gamma'$
and for all $j \in \{1,2,\dots,K\} \setminus \{y_t\}$,
$\ip{u_j}{x_t} \le - \gamma'$. In other words,
$(\rho(x_1), y_1), (\rho(x_2), y_2), \dots, (\rho(x_T), y_T)$ are
strongly linearly separable with margin $\gamma'$.
\end{proof}

>>>>

Using the lemma and the polynomial approximation theorems, we can prove the
mapping $\phi$ maps any a weakly separable data to a strongly separable data
set.

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T) \in \R^d \times \{1,2,\dots,K\}$
labeled examples that are weakly linearly separable with margin $\gamma > 0$
such that $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le 1$.
Let
\begin{align*}
r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1 \quad \text{and} \quad s = \left \lceil \log_2(1/\gamma) \right \rceil \\
\gamma_1 & = \frac{1}{2\sqrt{K} \left(376 \lceil \log_2(2m) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}} \\
\gamma_2 & = ??
\end{align*}
Then, the labeled examples $(\phi(x_1), y_1), (\phi(x_2), y_2), \dots,
(\phi(x_T), y_T)$ are strongly linearly separable with margin $\gamma' =
\max\{\gamma_1, \gamma_2\}$ and $\norm{\phi(x_1)}_{\ell_2},
\norm{\phi(x_2)}_{\ell_2}, \dots, \norm{\phi(x_T)}_{\ell_2} \le 2$.
\end{theorem}

\begin{proof}
First osberve that for any $t=1,2,\dots,T$,
$$
\norm{\phi(x_t)}_{\ell_2} = K_{\phi}(x_t,x_t) = \frac{1}{1 - \frac{1}{2} \norm{x_t}^2} \le 2 \; .
$$

@@

Fix any $i \in \{1,2,\dots,K\}$. Consider the $K-1$ vectors $(w_i - w_j)/2$ for
$j \in \{1,2,\dots,K\} \setminus \{i\}$. Note that the vectors have norm at most
$1$.
Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}
imply that there exists multivariate polynomials $p_i:\R^d \to \R$ and $q_i:\R^d
\to \R$ such that
\begin{align*}
\deg(p_i) & = \lceil \log_2(2K-2) \rceil \cdot \left\lceil \sqrt{\frac{2}{\gamma}} \right\rceil \; , \\
\deg(q_i) & = (2K-5/2) 2^m \\
& \qquad \cdot \left(2^{2s} r(K-1) (4s+2)^2 \right)^{(s+1/2)r(K-1)} \; .
\end{align*}
Futhermore, for all $t=1,2,\dots,T$, if $y_t = i$ then $p_i(x_t) \ge 1/2$, $q_i(x_t) \ge ???$ and if $y_t \neq i$ then $p_i(x_t) \le -1/2$,
$q_i(x) \le - ???$
and
\begin{align*}
\norm{p_i} & \le \left(188 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \; , \\
\norm{p_i} & \le ??? .
\end{align*}
By \autoref{lemma:norm-bound} there exists $c_i, c_i' \in \ell_2$ such that
$\ip{c_i}{\phi(x)} = p_i(x)$ and $\ip{c_i'}{\phi(x)} = q_i(x)$ and
\begin{align*}
\norm{c_i}_{\ell_2}
& \le \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \\
\norm{c'_i}_{\ell_2} & \le ??? \\
\end{align*}
Define vectors $u_i, u_i' \in \ell_2$ as
\begin{align*}
u_i & = \frac{c_i}{\sqrt{K} \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}} \; . \\
u_i' & = ???
\end{align*}

Then, $\norm{u_1}^2 + \norm{u_2}^2 + \dots + \norm{u_K}^2 \le 1$ and
$\norm{u_1'}^2 + \norm{u_2'}^2 + \dots + \norm{u_K'}^2 \le 1$.
Futhermore, for all $t=1,2,\dots,T$, $\ip{u_{y_t}}{x_t} \ge \gamma_1$ and
$\ip{u'_{y_t}}{x_t} \ge \gamma_2$
and for all $j \in \{1,2,\dots,K\} \setminus \{y_t\}$,
$\ip{u_j}{x_t} \le - \gamma_1$ and $\ip{u'_j}{x_t} \le - \gamma_2$. In other words,
$(\phi(x_1), y_1), (\phi(x_2), y_2), \dots, (\phi(x_T), y_T)$ are
strongly linearly separable with margin $\gamma_1$ and also strongly separable with
margin $\gamma_2$. Therefore, the examples are strongly separable with margin $\gamma' = \max\{\gamma_1, \gamma_2\}$.
\end{proof}
